---
title: "DESCRIBING RELATIONSHIPS"
subtitle: "1. Group Comparisons"
author: "Austin Hart"
institute: "American University"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, rladies, rladies-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  message=FALSE, warning=FALSE, eval = TRUE, echo = FALSE, 
  fig.align = 'center', dev='svg'
)
```

```{r results='hide'}
library(tidyverse)
library(magrittr)
library(kableExtra)

setwd("C:\\Users\\ahart\\Dropbox\\SIS 600 AH\\2021S 600\\content\\w07 associations\\GroupDiffs")

load('DCPS testing.RData')
```


# Group Comparison


Assessing the relationship between:  

- Nominal or binary exposure variable (X)  

- Continuous outcome variable (Y)  


Data used: "DCPS testing.RData"


---
# Summary stats revisited

.pull-left[
Math proficiency in 108 DC Public Schools based on 2017-18 PARCC assessment data.  

```{r sumtab}
  dcps %>%
    summarize(
      N = n(),
      Average = mean(ProfMath),
      Median = median(ProfMath)
    ) %>%
  kbl(digits = 1, caption = "Summary statistics",format = "html") %>%
  kable_paper(bootstrap_options = "striped", full_width = F, position = "center") 

```

]

.pull-right[
```{r boxsum, fig.width=3.5, fig.height=5, dpi=200}
  boxplot(dcps$ProfMath,
        main = "Math Proficiency, DCPS 2017",
        ylab = "% testing at grade level", 
        ylim = c(0,100), col="#69b3a2")
  abline(h = mean(dcps$ProfMath), col = 'red')
```
]

---
# Group comparison

> Does the distribution on the outcome measure (i.e. mean of Y) differ by groups?

- Calculate mean outcome for each group of X

- Compare and contrast: $\bar{Y}_H$ vs $\bar{Y}_M$ vs $\bar{Y}_E$


---
# Math proficiency by school level

.pull-left[

Does math proficiency differ by level of schooling in DC public schools?  


```{r sumtabgrp}
  dcps %>%
    group_by(SchType) %>%
    summarize(
      N = n(),
      Average = mean(ProfMath),
      Median = median(ProfMath)
    ) %>%
  kbl(digits = 1, caption = "Math Proficiency by Type", type = 'html') %>%
  kable_paper(bootstrap_options = "striped", full_width = F) 

```
]

.pull-right[
```{r boxsumgrp, fig.width=3.75, fig.height=5, dpi=250}
  boxplot(ProfMath ~ SchType, data = dcps,
        main = "Math Proficiency, DCPS 2017",
        xlab = NULL,
        ylab = "% testing at grade level", 
        ylim = c(0,100), col="#69b3a2",
        names=c("Elem.","Middle","High"))
```
]

---
class: inverse

# Group Comparison


- Assessing relationship between:  

  - Nominal or binary exposure variable (X)  

  - Continuous outcome variable (Y)  

- Process: calculate group means and compare!  
  
- More info  

  - R replication script  
  
  - RGuide  
  
  - Textbook chapter
